---
title             : "Effect of inhibitory control training on food-related action tendencies, liking and impulsive choice "
shorttitle        : "ICT EFFECTS ON FOOD-RELATED ACTION TENDENCIES, LIKING AND CHOICE"

author: 
  - name          : "Loukia Tzavella"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Cardiff University Brain Research Imaging Centre, CF24 4HQ, UK"
    email         : "tzavellal@cardiff.ac.uk"
  - name          : "Christopher D. Chambers"
    affiliation   : "1"
  - name          : "Natalia Lawrence"
    affiliation   : "2"
  - name          : "Katherine S. Button"
    affiliation   : "3"
  - name          : "Elizabeth Hart"
    affiliation   : "4"
  - name          : "Natalie Holmes"
    affiliation   : "4"
  - name          : "Kimberley Houghton"
    affiliation   : "4"
  - name          : "Nina Badkar"
    affiliation   : "2"
  - name          : "Ellie Macey"
    affiliation   : "2"
  - name          : "Amy-Jayne Braggins"
    affiliation   : "3"
  - name          : "Felicity Murray"
    affiliation   : "2"
  - name          : "Rachel C. Adams"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Cardiff University Brain Research Imaging Centre, CF24 4HQ, UK"
  - id            : "2"
    institution   : "School of Psychology, University of Exeter, EX4 4QG, UK"
  - id            : "3"
    institution   : "Department of Psychology, University of Bath, BS2 7AY, UK"
  - id            : "4"
    institution   : "School of Psychology, Cardiff University, CF10 3AT, UK"

authornote: |
  The research project was conducted as part of the GW4 Undergraduate Psychology Consortium 2017/2018 and was partially supported by the European Research Council (Consolidator 647893; C.D.C.). We also gratefully acknowledge Teaching Development Funding, from the faculty of Humanities and Social Sciences at the University of Bath for funding travel and room hire costs for the consortium meetings.


authornote: |
  [Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.]

  [Enter author note here.]

abstract: |
  [ADD ABSTRACT HERE]
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["zotero.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
header-includes   :
- \usepackage{tikz}
- \usepackage{xcolor}
  \hypersetup{
    colorlinks,
    linkcolor={red!60!blue},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
  }
  \urlstyle{same}
- \usepackage{multicol}
- \usepackage{enumitem}
  \setlist{nosep}
- \usepackage[font=footnotesize,labelfont=bf]{caption}
- \usepackage{amsmath}
- \usepackage{threeparttable}
- \usepackage{soul}
- \usepackage{caption}
- \usepackage{booktabs}
- \usepackage{float}
---

```{r setup, include = FALSE}
library("papaja")
```


# Introduction {#introduction}

\par

The recent rise in overweight and obesity rates can primarily be attributed to the over-consumption of energy-dense foods that are high in fat, sugar and salt content [@whoObesityOverweight2018]. One theoretical explanation for overeating has been provided by dual-process models which argue that behaviour is determined by the interaction of impulsive (*automatic*) and reflective (*controlled*) processes [@strack_reflective_2004; @kakoschke_combined_2015]. For example, over-consumption of unhealthy foods can be ascribed to heightened automatic biases for such foods, which can result in increased food intake if these automatic tendencies are not regulated via controlled processes [@kakoschke_effect_2017].$\color{red}{\text{provide example here – e.g. exposure to unhealthy foods in the environ (advertising)}}$ 

$\color{red}{\text{can trigger biased behavior and result in consumption for those who have limited self-control (either trait or state!}}$ 

\par

Theoretical frameworks, such as the dual process model, have led to the development of behaviour change interventions for unhealthy eating behaviours that target either automatic or controlled processing, such as approach bias modification and inhibitory control training respectively [see @kakoschke_approach_2017; @jonesCognitiveTrainingPotential2018 for recent reviews]. $\color{red}{\text{Need to take a step back here and explain ICT- Suggest discussion of inhibition, how it’s}}$ 
$\color{red}{\text{related to food-related behavior , training and associated effects on behavior. Discuss success in terms of }}$
$\color{red}{\text{findings and meta-analyses. Provide some key examples from the literature}}$


\par
$\color{red}\text{{Rachel's comment below:}}$
After discussing ICT and effects on behavior etc. I think we should discuss mechanism of devaluation… and then link to approach/ avoid-	Also need to provide some discussion of increased preferences for go foods… whilst appreciating the point that we do not have a cued-approach design (so there are some differences) but the logic / theory still applies


$\color{red}\text{{CUT bit:  }}$
The primary aim of the present study was to investigate the interaction between automatic and controlled processing in the context of inhibitory control training (ICT). Specifically, we tested whether food-specific ICT could influence individuals' automatic action tendencies towards unhealthy foods.

<!--\par
In the dual-process framework, unhealthy eating behaviours may be explained by a weak reflective system and/or a strong impulsive system [e.g., @lawrence_nucleus_2012; @nederkoorn_specificity_2012]. These systems are often in conflict; for example, when an individual has to decide whether or not to consume an unhealthy food, automatic attentional (e.g., attending to the cue) and motivational (e.g., approaching the food) processes would antagonize the controlled process of considering long-term goals such as losing weight when an individual has to decide on an action, that is to eat or not eat the food [@kakoschke_combined_2015]. -->

\par
This study focuses on an automatic process known as approach bias, which is the automatic action tendency to approach an appetitive (food) cue in the environment, rather than avoid it [@wiers_automatic_2013]. An approach bias has been demonstrated for a variety of energy-dense foods in both obese and healthy-weight individuals [@brignell_attentional_2009; @kemps_approach_2015; @kemps_implicit_2013; @veenstra_restrained_2010]. Interestingly, @kakoschke_combined_2015 found that approach bias alone did not predict increased intake of unhealthy foods, but it was the interaction between approach bias and inhibitory control that was the significant determinant of subsequent behaviour. The authors report that approach bias had the expected effect on food intake only for participants with low inhibitory control. An important component of controlled processing is inhibitory control, which refers to the ability to inhibit behaviours and impulses that are incompatible with higher-order goals, such as wanting to lose weight [@houben_too_2012], and encompasses several elements, such as response inhibition and cognitive flexibility [see @bartholdySystematicReviewRelationship2016]. Inhibitory control capacity is often measured via response inhibition paradigms, such as the go/no-go task [@dondersSpeedMentalProcesses1969; @newmanPassiveAvoidanceLearning1986] and stop-signal task [@loganAbilityInhibitSimple1984; @lappinUseDelayedSignal1966], and has been associated with unhealthy eating behaviours [e.g., @jasinska_impulsivity_2012;  @guerrieri_influence_2007; @hall_executive_2012]. For example, @nederkoorn_control_2010 showed that strong implicit preferences for snacks paired with low 'inhibitory control capacity' predicted weight gain over one year. Overall, there is evidence to suggest that both inhibitory control and motivational processes are important determinants of eating-related behaviour. 

\par
Complementary evidence for the role of automatic and controlled processes in the regulation of eating behaviours stems from the line of research dedicated to the development of health behaviour change interventions. Approach bias modification training is commonly delivered via an approach-avoidance task [AAT; @neumann_approach_2000; @rinck_approach_2007; @wiers_should_2013] and has been found to be effective in re-training approach bias for foods [@brockmeyer_approach_2015-1] and even reduce food intake in the laboratory [@schumacher_bias_2016; see @kakoschke_approach_2017 for review]. The AAT is assumed to capture automatic action tendencies when participants are instructed to respond to task-irrelevant feature such as the orientation (portrait or landscape) of the presented picture, by pulling or pushing a joystick [@wiers_automatic_2013]. The AAT can also pair actions with visual feedback, so that the picture gets bigger when participant pull the joystick towards them (zoom-in) and gets smaller when they push it away (zoom-out). Arm extension could indicate an approach response towards an appetitive food (object-reference) or an avoidance response when the food is pushed away from the body/self [@phaf_approach_2014] and thus visual feedback provides the self-reference attribute to the responses (e.g., object comes closer to one’s body). The 'zooming' feature disambiguates the mapping of responses to approach and avoidance actions, whereby pulling the joystick represents approach and pushing it reflects avoidance [@neumann_approach_2000]. In AAT training, contingencies between actions and stimuli are manipulated so that appetitive cues are associated with push actions (avoidance) and neutral items are paired with pull actions (approach). 

\par
In the context of controlled processes, ICT interventions involve cue-specific go/no-go or stop-signal tasks whereby participants are instructed to make a speeded choice response to appetitive stimuli such as foods or alcohol, but to withhold that response when a visual, or auditory, signal is presented. In go/no-go training, signal-stimulus mappings are manipulated so that appetitive cues (e.g., unhealthy foods) are consistently paired with a stop signal. Stopping to unhealthy foods has been shown to reduce food consumption [@adams_training_2017; @houben_chocolate_2015; @houben_training_2011; @lawrence_stopping_2015; @veling_using_2011; also see @allomDoesInhibitoryControl2016 for meta-analysis] and promote healthy food choices in the laboratory [@veling_stop_2013; @veling_training_2017]. ICT protocols have even been associated with increased weight loss [@lawrence_training_2015; @veling_targeting_2014]. A potential mechanism of action behind ICT effects on food consumption is stimulus devaluation [@veling_what_2017], whereby the evaluations of appetitive foods are reduced during training to facilitate performance when response inhibition is required [e.g., @chen_how_2016]. A possible explanation for this devaluation effect is provided by the Behaviour Stimulus Interaction (BSI) theory which posits that food stimuli are devalued when negative affect is induced to resolve the ongoing conflict between triggered approach reactions to appetitive foods and the need to inhibit responses towards those stimuli [@chen_how_2016; @veling_what_2017; @veling_when_2008]. When a food is devalued, the approach bias towards that cue is reduced and inhibition can successfully take place. It is possible that this reduction in approach bias can be observed after go/no-go training^[It should be noted that a link between affect and motivation is also included in the dual-process framework of (eating) behaviour]. 

\par
Theoretically, the effects of ICT could also be explained by hard-wired connections between Pavlovian appetitive/ aversive centres [dickinson_mechanisms_2014] and go/no-go responses [@mclaren_association_2016; @verbruggen_inhibitory_2014]. @verbruggen_inhibitory_2014 suggest that conditioned inhibitory control in ICT paradigms, such as go/no-go training, can have an influence not only on the evaluations of stimuli, but also their motivational value via links to the appetitive/aversive centres. For example, a stimulus consistently paired with stopping on signal trials would be devalued and the approach motivation for that stimulus could also be reduced via a hard-wired excitatory connection between the 'stop' system and the aversive centre [also see @dickinson_mechanisms_2014; @mclaren_association_2016].\hl{Evidence – eg. Guitar masip paper} Similarly, approach bias for foods consistently paired with go responses on no-signal trials could be increased via the link between the 'go' system and the appetitive centre.\hl{Let’s provide some evidence and study details}

\par
This study attempts to answer this question by employing a go/no-go training paradigm with unhealthy food stimuli and measuring automatic action tendencies via an AAT before and after training. We tested whether individuals would show reduced approach bias for the foods associated with response inhibition and/or increased approach bias for the foods paired with go responses after training. Consistent with previous ICT literature, the study also examined impulsive food choice and food liking as secondary training outcomes. The research study was conducted as part of the GW4 Undergraduate Psychology Consortium 2017/2018 and data were collected across Cardiff University, University of Exeter and University of Bath campuses. \hl{Last para needs to summarise the procedure and hypotheses}

# Hypotheses

NOTE: Rachel wants the section removed? "Remove section – should all be summarized in last para of the intro (don’t need learning hypoths in there though)"


All hypotheses described in this section are confirmatory and have been preregistered\footnote{Exact hypotheses from the preregistered protocol have been re-ordered according to outcomes for clarity. There were no deviations from the protocol for the hypotheses and corresponding statistical tests, with the exception of minor alterations regarding the supplementary frequentist statitics (see section \ref{ch2_prereg_analyses}).} on the Open Science Framework (https://osf.io/wav8p/). Effects of ICT (go/no-go training; see section \ref{ch2_gng}) on automatic action tendencies (see section \ref{ch2_aat}) and liking (see section \ref{ch2_food_ratings}) for unhealthy foods were investigated using change scores from pre-to post-training for both outcomes (H1, H3). The training condition was also expected to have an effect on food choice behaviour (H2; see section \ref{ch2_food_choice}). The study assessed contingency learning mechanisms for the training paradigm, as a manipulation check (H4).  

## Training effects on automatic action tendencies

The primary outcome measure in the study was the change in automatic action tendencies from pre-to post- ICT training for the foods associated with different conditions (go, no-go and control - see Figure \ref{fig:ch2_procedure}). Action tendencies were indirectly measured via the AAT and approach-avoidance bias scores were obtained by subtracting the median completion response times (RTs; for details see section \ref{ch2_measures_indices}) in avoid trials (push action) from the RTs in approach trials (pull action) at the participant level (correct responses only), for each training condition and then calculating the change from pre-to post-training. It was hypothesised that ICT training would lead to a reduction in approach bias for no-go foods and potential increase in approach bias for go foods compared to the control foods. \newline

\noindent H1. There would be evidence for an effect of training condition (go, no-go, control) on the change in approach-avoid bias scores from pre-to post-training.
\begin{itemize}
\item[H1a.] Participants would show a reduction in approach bias for no-go foods compared to the control foods, from pre-to post- training.
\item[H1b.] Participants would have increased approach bias towards go foods relative to the control foods, from pre-to post- training.
\end{itemize}

## Training effects on impulsive food choices

As a secondary outcome, the effects of ICT on impulsive food choices for unhealthy foods were tested by comparing the probabilities of choosing a food from each training condition. \newline

\noindent H2. It was expected that after ICT participants would show reduced impulsive choices for no-go foods and increased choices for go foods relative to control foods. 
\begin{itemize}
\item[H2a.] The probability of choosing a no-go food would be lower than the probability of choosing a control food.
\item[H2b.] The probability of choosing a go food would be higher than the probability of choosing a control food.
\end{itemize}

## Training effects on food evaluations

The mean change in food liking ratings from pre-to post-training was examined in order to test whether go/no-go (GNG) training led to the devaluation of no-go foods compared to control foods (H3a). It should be noted that this was not a positive control for training effectiveness, as the findings for stimulus devaluation outcomes remain controversial [see @jones_inhibitory_2016 for meta-analysis]. Nonetheless, stimulus devaluation in this study was treated both as a manipulation check for the employed training paradigm and a secondary outcome measure. The possibility that positive evaluations for go foods would be increased after training compared to the control foods was also considered and tested in a separate hypothesis (H3b). \newline

\noindent H3. There would be evidence for an effect of training condition (go, no-go, control) on the change in food liking from pre-to post-training.

\begin{itemize}
\item[H3a.] Participants would show reduced liking for no-go foods relative to the control foods, from pre-to post- training.
\item[H3b.] Participants would show increased liking for go foods relative to the control foods, from pre-to post- training.\footnote{This was only a hypothesis for secondary outcomes, as stimulus devaluation is exclusively defined as a reduction in positive evaluations for foods associated with no-go responses during training.}
\end{itemize}

## Contingency learning during training

Contingency learning was a more fundamental manipulation check for GNG training. ICT paradigms, such as the go/no-go training task, might lead to stimulus-response associations and learning can be observed in both reaction times and error rates [e.g., @lawrence_training_2015; see @best_instructed_2019]. The proportion of correct responses on signal trials (i.e., successful stops) and the mean reaction times from no-signal (go) trials were compared for specific training conditions, as stated in the hypotheses below. \newline

\noindent H4. Go/no-go training would result in contingency learning in terms of reaction times on no-signal trials and the percentage of successful inhibitions on signal trials. 
\begin{itemize}
\item[H4a.] The proportion of correct responses on signal trials would be greater for no-go foods compared to the control foods associated with a signal (control\textsubscript{nogo}).
\item[H4b.] Go reaction times would be faster for go foods compared to the control foods presented on no-signal trials (control\textsubscript{go}).
\end{itemize}


# Methods

# Results

# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
